Survival Analysis

### 2014-04-14 paper.Rnw
Finished background of introduction

Survival analysis has roots in actuarial sciences where life tables were used in calculating insurance rates centuries ago. Only within the last century has Survival Analysis become more popular.
% Background on Statistical Technique
% * Tries to answer "what proportion of the population will survive until x?"
% * started with actuarial tables
% * WWII - used for reliability analysis
% * Ed Kaplan & Paul Meier developed S.A. in cancer trials
%   + This is when S.A. took off, much interest
%   + Developed Kaplan-Meier survival estimator
% * Used in other fields (economics, engineering, sociology, ...)
% Objective of Paper
% * Guide one enoguh to be able to:
%   + design survival studies
%   + read survival/hazard curves
%   + estimate kaplan-meier
%   + use R to perform tests

@@

%%% References
% List all books, journal articles and web pages used.
%%% http://data.princeton.edu/wws509/notes/c7.pdf
%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
%%% http://www.jehps.net/juin2009/Aalenetal.pdf

>>>>

Survival analysis has roots in actuarial sciences where life tables were used in calculating insurance rates centuries ago. However, there has not been much advancement in the field until the last century. In World War II, military technology prowess did not keep up with its reliability until survival analysis was used to increase the reliability of equipment. Advances, however, were only limited to uses in the military and not embraced by the public. In the mid 1970s, Major Greenwood published a paper on cancer incidence and mortality by various types. Edward Kaplan and Paul Meier took notice of this study and developed what is known as the Kaplan-Meier estimator for the survival rate. After Kaplan and Meier published their landmark survival analysis paper, interest in the subject exploded and spread to other fields of study besides medical and engineering such as sociology and economics.
% Background on Statistical Technique
% * Tries to answer "what proportion of the population will survive until x?"
% * started with actuarial tables
% * WWII - used for reliability analysis
% * Ed Kaplan & Paul Meier developed S.A. in cancer trials
%   + This is when S.A. took off, much interest
%   + Developed Kaplan-Meier survival estimator
% * Used in other fields (economics, engineering, sociology, ...)
% Objective of Paper
% * Guide one enoguh to be able to:
%   + design survival studies
%   + read survival/hazard curves
%   + estimate kaplan-meier
%   + use R to perform tests

@@

%%% References
% List all books, journal articles and web pages used.
%%% http://data.princeton.edu/wws509/notes/c7.pdf
%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3059453/
%%% http://www.jehps.net/juin2009/Aalenetal.pdf
%%% www.worldofstatistics.org/files/2012/12/Stat2013.pptx
%%% http://books.google.com/books?id=1iI5rHtCT-4C&lpg=PA112&ots=2jMXda6Zjy&dq=%22major%20greenwood%22%20cancer&pg=PA112#v=onepage&q=%22major%20greenwood%22%20cancer&f=false
